---
title             : "SwehVd manual"
shorttitle        : "SwehVd manual"
date              : "`r format(Sys.time(), '%B %d, %Y')`"

author:
  - name: Anna Persson
    affiliation: '1'
    corresponding: yes
    address: 'Department of Swedish Language and Multilingualism, Stockholm University, SE-106 91 Stockholm, Sweden'
    email: 'anna.persson@su.se'
  - name: T. Florian Jaeger
    affiliation: '2,3'

affiliation:
  - id: '1'
    institution: Swedish Language and Multilingualism, Stockholm University
  - id: '2'
    institution: Brain and Cognitive Sciences, University of Rochester
  - id: '3'
    institution: Computer Science, University of Rochester
  
authornote: |
  We are grateful to ### ommitted for review ###
  
abstract: |
  XXX. All data and code for this study are shared via OSF, including the R markdown document that this article is generated from, and an R library that implements the models we present.
  
keywords: speech perception; vowels; distributional learning; computational model
wordcount: XXX

bibliography      : ["latex-stuff/library.bib", "latex-stuff/r-references.bib"]
link-citations: yes
csl: latex-stuff/apa-6th-edition.csl

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
numbersections    : true
mask              : no
draft             : no

header-includes:
 - \usepackage{animate}
 - \usepackage{amsmath}
 - \usepackage{tikz}
 - \usetikzlibrary{bayesnet}
 - \usepackage{booktabs}
 - \usepackage{siunitx}
 - \usepackage{soul}
 - \usepackage{tabto}
 - \usepackage{xcolor}
 - \usepackage{colortbl}
 - \usepackage{placeins}
 - \setstcolor{red}
 - \usepackage{sectsty}
 - \sectionfont{\color{black}}
 - \subsectionfont{\color{black}}
 - \subsubsectionfont{\color{black}}
 - \usepackage{setspace}\doublespacing
 - \usepackage{subfig}
 - \usepackage{lscape}
 - \usepackage{pdflscape}

documentclass     : "apa6"
classoption       : "man"
fontsize          : 11pt
output: 
  papaja::apa6_pdf:
    latex_engine: xelatex
    extra_dependencies: "subfig" 
    includes:
      in_header: latex-stuff/header.tex
always_allow_html: true
---

```{r libraries, include=FALSE, message=FALSE}
library(curl)               # Check availability of internet for install of remote libraries

# Install remotes package if necessary
if(!requireNamespace("remotes", quietly = TRUE)) if (has_internet()) install.packages("remotes")
# Install the stable development version from GitHub
if (has_internet()) remotes::install_github("crsh/papaja")
library(papaja)             # APA formatted ms

library(tidyverse)  # data wrangling and plotting
library(magrittr)

library(modelr)

library(plotly)     # for interactive HTML plots
library(mvtnorm)    # multivariate Gaussian distributions

if (has_internet()) remotes::install_github("hlplab/MVBeliefUpdatr")
library(MVBeliefUpdatr)  # for generating Ideal Observers

library(linguisticsdown)  # for inserting IPA symbols
library(kableExtra)       # for styling html format

library(assertthat)     # for error checking
library(rlang)
library(phonR)      # for vowel normalization functions
library(phonTools)  # for WattFabricius normalization function
```

```{r setup, include=FALSE}
library(knitr)

# Set knit defaults for code chunks
opts_chunk$set(
  dev = 'png', # default format of figures
  comment="",
  echo=FALSE, warning=FALSE, message=FALSE,
  cache=FALSE)

# some useful formatting functions for output of knitting
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

color_block = function(color) {
  function(x, options) sprintf('\\color{%s}\\begin{verbatim}%s\\end{verbatim}\\color{black}',
                               color, x)
}
knitr::knit_hooks$set(error = color_block('red'))
knitr::knit_hooks$set(warning = color_block('orange'))
```

```{r functions-and-pipes, include=FALSE}
source("../thesis/constants_thesis.R")

options(knitr.table.format = "latex")
color.transform <- "#F0F8FF"
color.centering <- "#FFE4E1"
color.centering.more  <- "#F0EAD6"

base.width = 3
base.height = 3
```

# To Do

  
  
```{r}

```
  
# Outline{#sec:outline}

This rmd contains information about the SwehVd database...



```{r}

```

# Talker information{#sec:talkers}

The following information is available on talker: age (at time of recording), gender and L1. Among the female L1 talkers, there were four bilinguals: with Spanish (1), French (1), Korean (1) and Persian (1) as their other L1.


# Production errors{#sec:errors}
In column 'Transcribed vowel', the correctly produced vowels are labelled 'intended', the mispronounced vowels are labelled 'unidentifiable'.



